Estimating duration in the presence of misclassification

### 2024-08-01 biology-data.tex
improve figure appearance
Christl comment on T+T data
Christl comment on representativeness

The CIS~\autocite{CIS} was set up in April 2020 to provide gold-standard measurements of the prevalence of SARS-CoV-2 in the community.
It was a longitudinal study with a representative sample of private households, \ie excluding settings such as care homes and student halls.
Initially, it was limited to England, but expanded to cover the whole of the UK in September 2020.
Enrolment was continuous until 31 Jan 2022, with data collected until 13 Mar 2023~\autocite{weiRisk}. 
The dataset is globally unique in providing a representative, longitudinal, and large-scale study across the pandemic.

The CIS had a household-based design, meaning that households were invited to participate in the study.
From the households that participated, all individuals aged 2 and over were invited to participate.
Once invited, an enrolment swab would be taken at the first visit followed by 4 further weekly visits (giving a total of 5 swabs on days 0, 7, 14, 21, 28 relative to enrolment) after which visits were monthly.
As well as a swab to test for SARS-CoV-2, participants were asked to complete a questionnaire, and some participants were asked to provide blood samples.
In reality, visits were often not on this precise schedule, and occasionally missed altogether.
A full description of the study can be found in the study protocol~\autocite{cisProtocol}.

@@

\begin{figure}
  \centering \includegraphics{biology-data/CIS-recruitment}
  \caption[CIS swab numbers]{%
    Number of swabs taken by the CIS up until Jan 2021.
    (A) Only enrolment (\ie first) swabs for each individual.
    (B) Total swabs.
    Note the differing y-axis scales.
    Data from \textcite{CIStechData}.
  }
  \label{biology-data:fig:CIS-recruitment}
\end{figure}
% \begin{figure}
%   \centering \includegraphics[width=\textwidth]{biology-data/CIS-num-tests}
%   \caption[Number of CIS tests]{%
%     Daily number of CIS tests conducted in the period considered in this thesis (Sep 2020 until Jan 2021).
%   }
%   \label{biology-data:fig:CIS-num-tests}
%   \todo[inline]{Cut this repetition of number of tests? Maybe highlight period of interest in previous plot}
% \end{figure}

@@

\begin{figure}
  \centering \includegraphics[width=\textwidth]{biology-data/CIS-positivity}
  \caption[CIS positivity]{%
    CIS positivity rate in the period considered in this thesis (Sep 2020 until Jan 2021).
    Dots show the unadjusted positivity rate (days with fewer than 100 tests are excluded for readability).
    Ribbon shows modelled 95\% CrI (Credible Interval) for the positivity, corrected for non-response bias and smoothed over time (using the methodology described in \cref{E-backcalc:sec:methods}).
  }
  \label{biology-data:fig:CIS-positivity}
\end{figure}

>>>>

The CIS~\autocite{CIS} was set up in April 2020 to provide gold-standard measurements of the prevalence of SARS-CoV-2 in the community.
It was a longitudinal study aiming for a representative sample of private households, \ie excluding settings such as care homes and student halls.
Initially, it was limited to England, but expanded to cover the whole of the UK in September 2020.
Enrolment was continuous until 31 Jan 2022, with data collected until 13 Mar 2023~\autocite{weiRisk}. 
The dataset is globally unique in providing a near-representative, longitudinal, and large-scale study across the pandemic.

The CIS had a household-based design, meaning that households were invited to participate in the study.
From the households that participated, all individuals aged 2 and over were invited to participate.
Once invited, an enrolment swab would be taken at the first visit followed by 4 further weekly visits (giving a total of 5 swabs on days 0, 7, 14, 21, 28 relative to enrolment) after which visits were monthly.
As well as a swab to test for SARS-CoV-2, participants were asked to complete a questionnaire, and some participants were asked to provide blood samples.
In reality, visits were often not on this precise schedule, and occasionally missed altogether.
Additionally, participants' test results from routine, community-based testing, known as NHS Test and Trace, were linked; I do not make use of this information within this thesis because I am evaluating the potential of the designed portion of CIS alone.
A full description of the study can be found in the study protocol~\autocite{cisProtocol}.

@@

\begin{figure}
  \makebox[\textwidth][c]{\includegraphics{biology-data/CIS-recruitment}}
  \caption[CIS swab numbers]{%
    Number of swabs taken by the CIS up until Jan 2021.
    (A) Only enrolment (\ie first) swabs for each individual.
    (B) Total swabs.
    Note the differing y-axis scales.
    Data from \textcite{CIStechData}.
  }
  \label{biology-data:fig:CIS-recruitment}
\end{figure}
% \begin{figure}
%   \makebox[\textwidth][c]{\includegraphics[width=\textwidth]{biology-data/CIS-num-tests}}
%   \caption[Number of CIS tests]{%
%     Daily number of CIS tests conducted in the period considered in this thesis (Sep 2020 until Jan 2021).
%   }
%   \label{biology-data:fig:CIS-num-tests}
%   \todo[inline]{Cut this repetition of number of tests? Maybe highlight period of interest in previous plot}
% \end{figure}

@@

\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=\textwidth]{biology-data/CIS-positivity}}
  \caption[CIS positivity]{%
    CIS positivity rate in the period considered in this thesis (Sep 2020 until Jan 2021).
    Dots show the unadjusted positivity rate (days with fewer than 100 tests are excluded for readability).
    Ribbon shows modelled 95\% CrI (Credible Interval) for the positivity, corrected for non-response bias and smoothed over time (using the methodology described in \cref{E-backcalc:sec:methods}).
  }
  \label{biology-data:fig:CIS-positivity}
\end{figure}

### 2024-08-01 introduction.tex
Christl comment on representativeness

The extraordinary impact of SARS-CoV-2 led the UK Government to fund two large prevalence surveys in representative samples of the general population.
These were designed to avoid the biases of community testing and observe the general population.
These surveys use RT-PCR (Real-Time Polymerase Chain Reaction, described in detail in \cref{E-biology-data:sec:PCR}) testing to detect the presence of the virus in a swab from an individual.
The first survey was the CIS (Coronavirus (COVID-19) Infection Survey), run by the ONS (Office for National Statistics)~\autocite{CIS,cisMethodsONS}.
The CIS was a longitudinal, household-based study, testing a large number of individuals weekly or monthly for SARS-CoV-2 infection (see \cref{E-intro:sec:cis} for a detailed description).

>>>>

The extraordinary impact of SARS-CoV-2 led the UK Government to fund two large prevalence surveys which aimed to recruit representative samples of the general population; such surveys are considered the gold-standard.
They invited individuals selected at random from the population; however, not all individuals selected responded to the invite introducing some non-representativeness (see \cref{E-biology-data:sec:MRP}).
These designs aim to avoid the biases of community testing and observe the general population.
These surveys use RT-PCR (Real-Time Polymerase Chain Reaction, described in detail in \cref{E-biology-data:sec:PCR}) testing to detect the presence of the virus in a swab from an individual.
The first survey was the CIS (Coronavirus (COVID-19) Infection Survey), run by the ONS (Office for National Statistics)~\autocite{CIS,cisMethodsONS}.
The CIS was a longitudinal, household-based study, testing a large number of individuals weekly or monthly for SARS-CoV-2 infection (see \cref{E-intro:sec:cis} for a detailed description).

### 2024-08-01 transmission.tex
improve figure appearance
Rob/Christl p123 comment
Christl comment on representativeness

In this chapter, I use the estimate of the distribution of the duration of RT-PCR positivity based on combining information from the ATACCC and CIS studies (see \cref{E-imperf-test:sec:conclusion}) to investigate the feasibility of estimating incidence and transmission of SARS-CoV-2 from only the CIS data (described in \cref{E-intro:sec:cis}).
I will show that the CIS data are sufficient, demonstrating the value of surveys of this type for pandemic surveillance.

@@

During November, incidence starts rising, with regional heterogeneity in the timing of when this occurs (see \cref{transmission:fig:backcalc-regions}).
The most explored explanation for this is the emergence of the Alpha variant, which was more transmissible than previous variants and emerged at different times across the country~\autocite{walkerTracking,daviesEstimated,lythgoeLineage}.
There is weak evidence that when incidence starts increasing is correlated with the timing of the Alpha variant's spread (see \cref{transmission:fig:backcalc-alpha}).
A second peak occurs in late December or early January in all regions; this second peak is larger than the first in all regions except the Northern three (North East, North West, and Yorkshire and the Humber).
\begin{figure}
    \includegraphics{transmission/backcalc-regions}
    \caption[Incidence estimated using the phenomenological approach by region]{%
        Incidence proportion for (A) all of England and (B) its component regions.
    }
    \label{transmission:fig:backcalc-regions}
\end{figure}
\begin{figure}
    \includegraphics{transmission/backcalc-alpha}
    \caption[Correlation of the emergence of Alpha and minimum incidence.]{%
        x-axis: time that S-gene target failure, a marker of the Alpha variant, is estimated to start increasing from baseline by region among CIS test results; from \textcite[table S3]{walkerTracking}.
        y-axis: posterior median and 95\% CrI of the date of minimum incidence in November or early December in each region.
        Linear regression line shown.
    }
    \label{transmission:fig:backcalc-alpha}
\end{figure}

The incidence proportion was highest among the 12--16 and 17--25 age groups throughout the period (see \cref{transmission:fig:backcalc-ages}).
Except the 17--25 age group, all age groups show similar trends (\ie parallel lines on a log scale), meaning there was insufficient information in the data to overcome the prior which favoured this situation.
The incidence proportion in the 17--25 age group decreases faster than other age groups during November.
There are many plausible explanations for this effect.
The higher incidence in this age group prior to November means it could be an immunity effect.
Alternatively, it could be a consequence of schools remaining open in the November lockdown; 20--24 year olds rarely interact with school-aged children (see \cref{E-SEIR:fig:age-contacts}).
Further work is required to determine causality.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{transmission/backcalc-ages}}
    \caption[Incidence estimated using the phenomenological approach by age group]{%
        Incidence proportion for each age group, on both natural and log scales.
    }
    \label{transmission:fig:backcalc-ages}
\end{figure}

An edge effect, a rapid rise in incidence between days one and two, is visible (see \cref{transmission:fig:backcalc-start-effect}).
This is due to the incorrect assumption that incidence is constant in the past.
Changing this assumption (\eg to exponential growth) modifies the estimated value on the first day, but the rest of the series is insensitive to this change (not shown).
Therefore, the first day's estimate is unreliable, and has been excluded from all other results.
However, the rest of the series is robust to this assumption.
\begin{figure}
    \centering \includegraphics{transmission/backcalc-start-effect}
    \caption[Edge effects in the phenomenological approach]{%
        First seven days of incidence proportion estimate for the whole of England, highlighting the edge effect.
    }
    \label{transmission:fig:backcalc-start-effect}
\end{figure}

@@

In this section, I apply the general mechanistic transmission modelling framework from \cref{E-SEIR} to the CIS prevalence data.
The application of this framework follows \textcite{birrellRealtime}.
The largest modification from the generic framework presented in \cref{E-SEIR} is the parameterization of the WAIFW matrix (\cref{SEIR:sec:WAIFW-application}).
A novel approach (\cref{SEIR:sec:observation}) links this transmission model to the CIS prevalence data; the approach is generalisable to other studies collecting representative prevalence data, \eg REACT.
In the mechanistic modelling literature, this part of the model is referred to as the observation model, however, in the terminology of \cref{E-inc-prev}, it includes both the prevalence process and the observation process.

@@

Next, I solved the ODE system in \cref{SEIR:eq:fullODEs} for each parameter set.
All other aspects of the ODEs, including the data-informed contact matrices, $\matr{\kappa}(t)$, are as previously described in this section.
Population sizes were those of the East of England, and the time period the same as the main analysis.
This means that the simulations created realistic incidence and prevalence curves (see \cref{SEIR:fig:sim-data}).
In particular, the October half-term school holiday is visible, especially in the school-age groups and the November lockdown is visible to some extent in all age groups.
\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/sim/data}}
    \caption[Simulated data]{%
        (A) Incidence and (B) prevalence for the 100 simulated data sets as a proportion of each stratum.
        Prevalence shown is $\bar{p}_{kat}$, the true population prevalence on each day, excluding sampling noise.
        Each line represents one simulation.
    }
    \label{SEIR:fig:sim-data}
\end{figure}

@@

The majority of the confidence intervals contain the nominal probability (see \cref{SEIR:fig:sim-coverage}).
There is a tendency for the coverage to be slightly lower than the nominal probability, suggesting that the credible intervals are slightly too narrow.
This is a common issue with random walk MCMC (see \cref{E-MCMC:sec:improving}).
The same issue arises with the coverage for derived quantities incidence and prevalence (see \cref{SEIR:fig:sim-inc-prev}).
\begin{figure}
    \includegraphics{SEIR/sim/coverage}
    \caption[Coverage of simulation study]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for each parameter; see text for full definition.
        Vertical lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-coverage}
\end{figure}
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics[width=0.9\paperwidth]{SEIR/sim/predictive_coverage}}
    \caption[Coverage of simulation study (derived quantities)]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for (A) incidence and (B) prevalence; see main text for full definition.
        Horizontal lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-inc-prev}
\end{figure}

@@

For $\vec{\pi}$ and $\rho$, the correlation between the posterior median and the generative value is weak (see \cref{SEIR:fig:true-vs-posterior}).
This suggests that these parameters are not practically identifiable in this setting.
The posteriors align very closely with the priors (not shown) which reinforces this finding.
The lack of identifiability for $\vec{\pi}$ aligns with previous work showing this parameter is structurally unidentifiable, \ie not identifiable even in the limit of infinite data~\autocite{dankwaStructural}.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{SEIR/sim/true_vs_posterior}}
    \caption[True vs posterior parameter values]{%
        True vs posterior median parameter values for each parameter in each simulation.
        The black line shows the line $y = x$ which the dots would lie on if the parameter is fully recovered.
        The blue line is a linear regression line.
    }
    \label{SEIR:fig:true-vs-posterior}
\end{figure}

@@

The 12--16 age group (roughly corresponding to compulsory secondary school education) is estimated to have the highest incidence (see \cref{SEIR:fig:incidence}).
Incidence is smooth within each week but shows discontinuities between weeks.
This is because the parameters within $\matr{\beta}(t)$ are constant within each week.
A constant $\matr{\beta}(t)$ would lead to exponential growth following a period of transitory dynamics, although the weekly switching may mean that these transitory dynamics never end, especially when incidence is high~\autocite{rhodesConvergence}.
The estimated timing of the peak in incidence varies by region, with the most common time being late December 2020 (see \cref{SEIR:fig:peak-incidence}).
In Yorkshire and the Humber, the October peak is estimated to be greater than the winter peak; for North West there is a significant probability on either being larger.
\begin{figure}
    \includegraphics{SEIR/CIS/incidence}
    \caption[Posterior estimates of incidence]{%
        Posterior estimates (95\% CrI) of incidence proportion by region and age.
    }
    \label{SEIR:fig:incidence}
\end{figure}
\begin{figure}
    \includegraphics{SEIR/CIS/p_peak}
    \caption[Posterior estimates of peak incidence timing]{%
        Posterior probability of the peak incidence proportion for each region and age combination occurring on each day.
    }
    \label{SEIR:fig:peak-incidence}
\end{figure}

@@

\begin{figure}
    \includegraphics{SEIR/CIS/beta_walk}
    \caption[Posterior estimates of $\beta_w$]{%
        Posterior estimates $\beta_w$ for each week.
        Shown are ``violins'' where the width is proportional to the kernel-density estimated posterior density at that point.
        Values trend downwards until early December when an increase starts.
    }
    \label{SEIR:fig:beta-walk}
\end{figure}

@@

Overall, the model fits the data well (see \cref{SEIR:fig:prev,SEIR:fig:prev-appendix}).
There are a few features of the data which are not captured by the model and the posterior estimate appears to be a compromise with the prevalence too low in one age group and too high in another.
For example, in the North East region, the 12--16 age group has its prevalence overestimated in October but the 17--25 age group has its prevalence underestimated.
\begin{figure}
    % \thisfloatpagestyle{empty}
    % \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_main}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit]{%
        Goodness-of-fit to prevalence data by region and age for selected regions (see \cref{SEIR:fig:prev-appendix} for others).
        The central ribbon and line show the posterior median and 95\% CrI estimate for the proportion of the relevant stratum RT-PCR positive on each day.
        The lighter outer ribbon is the 95\% posterior predictive interval for the proportion of test results that are positive each week.
        The dots show the observed prevalence in CIS, aggregated by week.
        For a well-calibrated model, 95\% of the dots would be within the outer ribbon.
    }
    \label{SEIR:fig:prev}
\end{figure}

@@

\Cref{transmission:fig:compare-regions} compares the regional incidence estimates between the two models in this chapter.
Generally, the estimates agree with the SEIR model having much narrower credible intervals.
The largest discrepancy is that the phenomenological approach model estimates a second peak in mid-January in several regions (most clearly in London, North West, and the East of England).
However, except in London, the credible intervals still overlap.
This is a specific case of a larger pattern: the SEIR model produces smoother estimates of incidence than the phenomenological model.
\begin{figure}
    \centering \includegraphics{transmission/compare-regions}
    \caption[Comparing each models estimate of regional incidence.]{%
        Incidence proportion in England and each region estimated by the two models discussed in this chapter.
        Posterior median and 95\% CrI shown.
        For the SEIR model, newly RT-PCR positive (entrances into the P1 compartment), rather than incidence (entrances into the E compartment) are shown for comparability with the phenomenological results.
    }
    \label{transmission:fig:compare-regions}
\end{figure}

@@

Many of these issues are clearly demonstrated by the estimates for the North East (see \cref{transmission:fig:compare-NE}).
I previously observed the quality of the SEIR model's fit is variable between different age groups (see \cref{SEIR:fig:prev,SEIR:fig:prev-appendix}).
This is because the model is not flexible enough to capture the true age distribution of infections.
The overestimation of prevalence in the 12--16 age group and underestimation in the 17--25 age group is reflected when comparing incidence estimates to those from the phenomenological model, which have reduced bias.
\begin{figure}
    \centering \includegraphics{transmission/compare-NE}
    \caption[Comparing each models estimate of North East incidence proportion by age.]{%
        Incidence proportion in each age group in the North East estimated by the two models discussed in this chapter.
        Posterior median and 95\% CrI shown.
        For the SEIR model, newly RT-PCR positive (entrances into the P1 compartment), rather than incidence (entrances into the E compartment) are shown for comparability with the phenomenological results.
    }
    \label{transmission:fig:compare-NE}
\end{figure}

>>>>

In this chapter, I use the estimate of the distribution of the duration of RT-PCR positivity based on combining information from the ATACCC and CIS studies (see \cref{E-imperf-test:sec:conclusion}) to investigate the feasibility of estimating incidence and transmission of SARS-CoV-2 from only the CIS data (described in \cref{E-intro:sec:cis}).
I will show that this is feasible, and the estimates have narrow enough credible intervals, to be useful for informing the response to a pandemic.

@@

During November, incidence starts rising, with regional heterogeneity in the timing of when this occurs (see \cref{transmission:fig:backcalc-regions}).
The most explored explanation for this is the emergence of the Alpha variant, which was more transmissible than previous variants and emerged at different times across the country~\autocite{walkerTracking,daviesEstimated,lythgoeLineage}.
There is weak evidence that when incidence starts increasing is correlated with the timing of the Alpha variant's spread (see \cref{transmission:fig:backcalc-alpha}).
A second peak occurs in late December or early January in all regions; this second peak is larger than the first in all regions except the Northern three (North East, North West, and Yorkshire and the Humber).
\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics{transmission/backcalc-regions}}
    \caption[Incidence estimated using the phenomenological approach by region]{%
        Incidence proportion for (A) all of England and (B) its component regions.
    }
    \label{transmission:fig:backcalc-regions}
\end{figure}
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{transmission/backcalc-alpha}}
    \caption[Correlation of the emergence of Alpha and minimum incidence.]{%
        x-axis: time that S-gene target failure, a marker of the Alpha variant, is estimated to start increasing from baseline by region among CIS test results; from \textcite[table S3]{walkerTracking}.
        y-axis: posterior median and 95\% CrI of the date of minimum incidence in November or early December in each region.
        Linear regression line shown.
    }
    \label{transmission:fig:backcalc-alpha}
\end{figure}

The incidence proportion was highest among the 12--16 and 17--25 age groups throughout the period (see \cref{transmission:fig:backcalc-ages}).
Except the 17--25 age group, all age groups show similar trends (\ie parallel lines on a log scale), meaning there was insufficient information in the data to overcome the prior which favoured this situation.
The incidence proportion in the 17--25 age group decreases faster than other age groups during November.
There are many plausible explanations for this effect.
The higher incidence in this age group prior to November means it could be an immunity effect.
Alternatively, it could be a consequence of schools remaining open in the November lockdown; 20--24 year olds rarely interact with school-aged children (see \cref{E-SEIR:fig:age-contacts}).
Further work is required to determine causality.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{transmission/backcalc-ages}}
    \caption[Incidence estimated using the phenomenological approach by age group]{%
        Incidence proportion for each age group, on both natural and log scales.
    }
    \label{transmission:fig:backcalc-ages}
\end{figure}

An edge effect, a rapid rise in incidence between days one and two, is visible (see \cref{transmission:fig:backcalc-start-effect}).
This is due to the incorrect assumption that incidence is constant in the past.
Changing this assumption (\eg to exponential growth) modifies the estimated value on the first day, but the rest of the series is insensitive to this change (not shown).
Therefore, the first day's estimate is unreliable, and has been excluded from all other results.
However, the rest of the series is robust to this assumption.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{transmission/backcalc-start-effect}}
    \caption[Edge effects in the phenomenological approach]{%
        First seven days of incidence proportion estimate for the whole of England, highlighting the edge effect.
    }
    \label{transmission:fig:backcalc-start-effect}
\end{figure}

@@

In this section, I apply the general mechanistic transmission modelling framework from \cref{E-SEIR} to the CIS prevalence data.
The application of this framework follows \textcite{birrellRealtime}.
The largest modification from the generic framework presented in \cref{E-SEIR} is the parameterization of the WAIFW matrix (\cref{SEIR:sec:WAIFW-application}).
A novel approach (\cref{SEIR:sec:observation}) links this transmission model to the CIS prevalence data; the approach is generalisable to other gold-standard prevalence studies, \eg REACT.
In the mechanistic modelling literature, this part of the model is referred to as the observation model, however, in the terminology of \cref{E-inc-prev}, it includes both the prevalence process and the observation process.

@@

Next, I solved the ODE system in \cref{SEIR:eq:fullODEs} for each parameter set.
All other aspects of the ODEs, including the data-informed contact matrices, $\matr{\kappa}(t)$, are as previously described in this section.
Population sizes were those of the East of England, and the time period the same as the main analysis.
This means that the simulations created realistic incidence and prevalence curves (see \cref{SEIR:fig:sim-data}).
In particular, the October half-term school holiday is visible, especially in the school-age groups and the November lockdown is visible to some extent in all age groups.
\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics{SEIR/sim/data}}
    \caption[Simulated data]{%
        (A) Incidence and (B) prevalence for the 100 simulated data sets as a proportion of each stratum.
        Prevalence shown is $\bar{p}_{kat}$, the true population prevalence on each day, excluding sampling noise.
        Each line represents one simulation.
    }
    \label{SEIR:fig:sim-data}
\end{figure}

@@

The majority of the confidence intervals contain the nominal probability (see \cref{SEIR:fig:sim-coverage}).
There is a tendency for the coverage to be slightly lower than the nominal probability, suggesting that the credible intervals are slightly too narrow.
This is a common issue with random walk MCMC (see \cref{E-MCMC:sec:improving}).
The same issue arises with the coverage for derived quantities incidence and prevalence (see \cref{SEIR:fig:sim-inc-prev}).
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{SEIR/sim/coverage}}
    \caption[Coverage of simulation study]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for each parameter; see text for full definition.
        Vertical lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-coverage}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics{SEIR/sim/predictive_coverage}}
    \caption[Coverage of simulation study (derived quantities)]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for (A) incidence and (B) prevalence; see main text for full definition.
        Horizontal lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-inc-prev}
\end{figure}

@@

For $\vec{\pi}$ and $\rho$, the correlation between the posterior median and the generative value is weak (see \cref{SEIR:fig:true-vs-posterior}).
This suggests that these parameters are not practically identifiable in this setting.
The posteriors align very closely with the priors (not shown) which reinforces this finding.
The lack of identifiability for $\vec{\pi}$ aligns with previous work showing this parameter is structurally unidentifiable, \ie not identifiable even in the limit of infinite data~\autocite{dankwaStructural}.
\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics{SEIR/sim/true_vs_posterior}}
    \caption[True vs posterior parameter values]{%
        True vs posterior median parameter values for each parameter in each simulation.
        The black line shows the line $y = x$ which the dots would lie on if the parameter is fully recovered.
        The blue line is a linear regression line.
    }
    \label{SEIR:fig:true-vs-posterior}
\end{figure}

@@

The 12--16 age group (roughly corresponding to compulsory secondary school education) is estimated to have the highest incidence (see \cref{SEIR:fig:incidence}).
Incidence is smooth within each week but shows discontinuities between weeks.
This is because the parameters within $\matr{\beta}(t)$ are constant within each week.
A constant $\matr{\beta}(t)$ would lead to exponential growth following a period of transitory dynamics, although the weekly switching may mean that these transitory dynamics never end, especially when incidence is high~\autocite{rhodesConvergence}.
The estimated timing of the peak in incidence varies by region, with the most common time being late December 2020 (see \cref{SEIR:fig:peak-incidence}).
In Yorkshire and the Humber, the October peak is estimated to be greater than the winter peak; for North West there is a significant probability on either being larger.
\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics{SEIR/CIS/incidence}}
    \caption[Posterior estimates of incidence]{%
        Posterior estimates (95\% CrI) of incidence proportion by region and age.
    }
    \label{SEIR:fig:incidence}
\end{figure}
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{SEIR/CIS/p_peak}}
    \caption[Posterior estimates of peak incidence timing]{%
        Posterior probability of the peak incidence proportion for each region and age combination occurring on each day.
    }
    \label{SEIR:fig:peak-incidence}
\end{figure}

@@

\begin{figure}
    \thisfloatpagestyle{empty}
    \makebox[\textwidth][c]{\includegraphics{SEIR/CIS/beta_walk}}
    \caption[Posterior estimates of $\beta_w$]{%
        Posterior estimates $\beta_w$ for each week.
        Shown are ``violins'' where the width is proportional to the kernel-density estimated posterior density at that point.
        Values trend downwards until early December when an increase starts.
    }
    \label{SEIR:fig:beta-walk}
\end{figure}

@@

Overall, the model fits the data well (see \cref{SEIR:fig:prev,SEIR:fig:prev-appendix}).
There are a few features of the data which are not captured by the model and the posterior estimate appears to be a compromise with the prevalence too low in one age group and too high in another.
For example, in the North East region, the 12--16 age group has its prevalence overestimated in October but the 17--25 age group has its prevalence underestimated.
\begin{figure}
    \thisfloatpagestyle{empty}
    % \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics{SEIR/CIS/prev_main}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit]{%
        Goodness-of-fit to prevalence data by region and age for selected regions (see \cref{SEIR:fig:prev-appendix} for others).
        The central ribbon and line show the posterior median and 95\% CrI estimate for the proportion of the relevant stratum RT-PCR positive on each day.
        The lighter outer ribbon is the 95\% posterior predictive interval for the proportion of test results that are positive each week.
        The dots show the observed prevalence in CIS, aggregated by week.
        For a well-calibrated model, 95\% of the dots would be within the outer ribbon.
    }
    \label{SEIR:fig:prev}
\end{figure}

@@

\Cref{transmission:fig:compare-regions} compares the regional incidence estimates between the two models in this chapter.
Generally, the estimates agree with the SEIR model having much narrower credible intervals.
The largest discrepancy is that the phenomenological approach model estimates a second peak in mid-January in several regions (most clearly in London, North West, and the East of England).
However, except in London, the credible intervals still overlap.
This is a specific case of a larger pattern: the SEIR model produces smoother estimates of incidence than the phenomenological model.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{transmission/compare-regions}}
    \caption[Comparing each models estimate of regional incidence.]{%
        Incidence proportion in England and each region estimated by the two models discussed in this chapter.
        Posterior median and 95\% CrI shown.
        For the SEIR model, newly RT-PCR positive (entrances into the P1 compartment), rather than incidence (entrances into the E compartment) are shown for comparability with the phenomenological results.
    }
    \label{transmission:fig:compare-regions}
\end{figure}

@@

Many of these issues are clearly demonstrated by the estimates for the North East (see \cref{transmission:fig:compare-NE}).
I previously observed the quality of the SEIR model's fit is variable between different age groups (see \cref{SEIR:fig:prev,SEIR:fig:prev-appendix}).
This is because the model is not flexible enough to capture the true age distribution of infections.
The overestimation of prevalence in the 12--16 age group and underestimation in the 17--25 age group is reflected when comparing incidence estimates to those from the phenomenological model, which have reduced bias.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics{transmission/compare-NE}}
    \caption[Comparing each models estimate of North East incidence proportion by age.]{%
        Incidence proportion in each age group in the North East estimated by the two models discussed in this chapter.
        Posterior median and 95\% CrI shown.
        For the SEIR model, newly RT-PCR positive (entrances into the P1 compartment), rather than incidence (entrances into the E compartment) are shown for comparability with the phenomenological results.
    }
    \label{transmission:fig:compare-NE}
\end{figure}
